\documentclass[conference]{IEEEtran}
\usepackage{cite}
\usepackage{amsmath,amssymb,amsfonts}
\usepackage{algorithmic}
\usepackage{graphicx}
\usepackage{textcomp}
\usepackage{xcolor}
\usepackage{listings}
\usepackage{hyperref}
\usepackage{tikz}
\usepackage{pgfplots}
\pgfplotsset{compat=1.17}
\usetikzlibrary{shapes,arrows,positioning,calc}

% Define a soft background color for the code blocks
\definecolor{codebg}{RGB}{245,245,248}

% Code listing style configuration
\lstset{
  language=C++,
  backgroundcolor=\color{codebg}, 
  basicstyle=\ttfamily\scriptsize, 
  keywordstyle=\color{blue},
  commentstyle=\color{green!60!black},
  stringstyle=\color{purple},
  breaklines=true,
  showstringspaces=false,
  frame=single,
  rulecolor=\color{gray!50}, 
  xleftmargin=0.5em,  
  xrightmargin=0.5em, 
  captionpos=b
}

\begin{document}

\title{A Morse Code-based Optical Communication System}
%\title{Adaptive Optical Telemetry: Evolving from Visible Light Baselines to Infrared Stepwise Derivative Thresholding and Goertzel Filtering}

\author{\IEEEauthorblockN{Kartik Lahoti, Puni Aditya and G V V Sharma}
\IEEEauthorblockA{\textit{Department of Electrical Engineering} \\
\textit{IIT Hyderabad}}
}

\maketitle

\begin{abstract}
%Optical wireless communication offers a license-free alternative to Radio Frequency (RF) systems. However, simple visible light links utilizing Light Dependent Resistors (LDRs) face significant challenges regarding ambient light interference and hardware latency. 
This paper outlines a comprehensive multi-stage development of a Morse Code On-Off Keying (OOK) system. We first establish a baseline Visible Light Communication (VLC) link to demonstrate the fundamental principles of communication over an optical link.  
%, documenting a maximum range of 27 cm and identifying the "Iceberg Effect" of signal truncation. 
Subsequently, 
%Phase II 
we introduce 
%the transition to the 
Infrared (IR) 
communication to extend the range and explore more advanced techniques to improve the system performance. 
This is done by using the Arduino as the transceiver platform.  Our communication scheme is successfully demonstrated for a range of up to 2 meters.
%spectrum coupled with a Stepwise Derivative Method (SDM) for adaptive thresholding, extending the range to 1.57 meters indoors. Finally, Phase III addresses the critical limitation of direct solar interference by implementing a fully software-defined 2~kHz carrier wave and a digital Goertzel filter, successfully rejecting DC solar saturation without requiring complex commercial hardware filters.
\end{abstract}

\begin{IEEEkeywords}
Visible Light Communication (VLC), Infrared (IR), Stepwise Derivative Method (SDM), Goertzel Algorithm, On-Off Keying (OOK), Adaptive Thresholding.
\end{IEEEkeywords}

\section{Introduction}
Wireless optical communication is highly dependent on clear Line-of-Sight (LoS) and a high Signal-to-Noise Ratio (SNR) \cite{elgala2011, ghassemlooy2012}. In low-cost embedded systems, a standard Light Emitting Diode (LED) paired with a Light Dependent Resistor (LDR) is commonly used to demonstrate data transfer. However, ambient environmental lighting introduces a severe DC bias into the receiver, and LDRs exhibit significant photo-conductive latency \cite{sewaiwar2015}. 
The novelty of this paper lies in its evolutionary approach to solving physical layer limitations, building upon foundational On-Off Keying (OOK) optical paradigms \cite{komine2004}. We first establish a baseline VLC system to document hardware-induced signal degradation, known as the "Iceberg Effect." Following this, we introduce an advanced Infrared (IR) iteration utilizing the Stepwise Derivative Method (SDM)—a software-defined algorithm that tracks the environmental noise floor \cite{green2005}. Finally, to overcome severe outdoor solar interference \cite{rajagopal2012}, we introduce a Goertzel algorithm-based digital signal processing (DSP) approach, effectively turning a low-cost microcontroller into a software-defined radio for optical frequencies.

\subsection{Related Works and Novelty}
Standard commercial IR systems avoid ambient noise by utilizing 38kHz Carrier Waves and specialized TSOP receiver modules. While effective, these modules act as "black boxes" that strip the analog signal data away from the microcontroller, preventing custom low-level signal analysis. Conversely, standard unmodulated photodiode circuits use fixed analog thresholds that fail the moment room lighting changes. Our proposed SDM and Goertzel implementations bridge this gap, retaining raw analog sampling while applying dynamic digital filters to reject ambient DC bias. The related block diagram is available in 
    Fig. \ref{fig:block}
\begin{figure}[htbp]
    \centering
    \resizebox{\columnwidth}{!}{
    \begin{tikzpicture}[node distance=1.2cm, auto, >=stealth]
        \tikzstyle{block} = [rectangle, draw, fill=blue!10, text width=6em, text centered, rounded corners, minimum height=3em]
        \tikzstyle{line} = [draw, -latex', thick]
        
        % TX Nodes
        \node [block] (mcu1) {TX MCU\\(Morse Encoder)};
        \node [block, right=of mcu1] (led) {LED/IR LED\\(OOK Driver)};
        
        % Channel
        \node [coordinate, right=2.5cm of led] (ch1) {}; 
        \node [coordinate, below=1.5cm of ch1] (ch2) {};
        
        % RX Nodes
        \node [block, below=1.5cm of led] (sensor) {LDR/PD\\(Voltage Divider)};
        \node [block, left=of sensor] (adc) {ADC\\(Sampling)};
        \node [block, below=of adc] (sdm) {SDM/Goertzel\\(Digital Filter)};
        \node [block, right=of sdm] (decoder) {Symbol\\Decoder};
        \node [block, right=of decoder] (disp) {7-Segment\\Display};

        % Paths
        \path [line] (mcu1) -- node {Digital} (led);
        \path [line, dashed, red] (led) -- node[above] {Optical Channel} (ch1) -- (ch2) -- (sensor);
        \path [line] (sensor) -- node[above] {Analog} (adc);
        \path [line] (adc) -- (sdm);
        \path [line] (sdm) -- node[above] {Pulses} (decoder);
        \path [line] (decoder) -- node[above] {Char} (disp);
    \end{tikzpicture}
    }
    \caption{System Block Diagram detailing the cross-layer flow from hardware transmission to the software-defined filtering and hardware output.}
    \label{fig:block}
\end{figure}
%\section{Hardware Configuration}
To ensure reproducibility, the system was tested using standard off-the-shelf components. The development was split into two hardware profiles as detailed in Table \ref{tab:components}.
\begin{table}[htbp]
\caption{List of Components}
\label{tab:components}
\centering
\begin{tabular}{|l|l|}
\hline
\textbf{Setup} & \textbf{Components} \\ \hline
\textbf{Common} & 2$\times$ Arduino UNO, \\
&  2$\times$ Breadboards, \\
& Jumper Cables \\ \hline
\textbf{Visible} & 1$\times$ Red LED ($5\text{mm}$), \\
& 1$\times$ LDR ($5\text{mm}$) \\ \hline
\textbf{IR} & 1$\times$ 940-nm IR LED, \\
& 1$\times$ IR Photodiode (Reverse Bias) \\ \hline
\end{tabular}
\end{table}

\subsection{Baud Rate and Telemetry Applicability}
Morse code, utilizing OOK, possesses a fundamentally low baud rate compared to modern digital modulation schemes like UART or QAM. However, for specialized telemetry applications—such as underwater optical communication or low-cost sensor nodes—signal penetration and robustness take precedence over raw speed. The widened tolerance windows used in this paper's decoding logic demonstrate how low baud rate systems can reliably reconstruct highly degraded signals \cite{uysal2015}.

\subsection{Baseband Morse Encoding}
To establish a quantitative baseline for On-Off Keying (OOK) modulation, the system utilizes the standard international Morse code protocol. In this telemetry model, the transmitter's digital output pin toggles between a HIGH (active) and LOW (idle) state to construct characters. 

The protocol defines all signal durations relative to a single base time unit ($T_u$), which in our experimental setup is set to $200\text{ ms}$. The strict timing rules governing this variable-length baseband modulation are detailed in Table \ref{tab:morse_timing}. 

\begin{table}[htbp]
\caption{Standard Morse Protocol Timings ($T_u = 200\text{ ms}$)}
\label{tab:morse_timing}
\centering
\begin{tabular}{|l|c|l|}
\hline
\textbf{Protocol Element} & \textbf{Logic State} & \textbf{Duration} \\ \hline
Dot (Dit)                 & HIGH (ON)            & 1 $T_u$ ($200\text{ ms}$) \\ \hline
Dash (Dah)                & HIGH (ON)            & 3 $T_u$ ($600\text{ ms}$) \\ \hline
Intra-character gap       & LOW (OFF)            & 1 $T_u$ ($200\text{ ms}$) \\ \hline
Inter-character gap       & LOW (OFF)            & 3 $T_u$ ($600\text{ ms}$) \\ \hline
Inter-word gap            & LOW (OFF)            & 7 $T_u$ ($1400\text{ ms}$) \\ \hline
\end{tabular}
\end{table}

This baseband square wave serves as the fundamental logic envelope for all subsequent experimental phases. A visual representation of these timing constraints is illustrated in Fig. \ref{fig:morse_an} for the transmission of the word "AN".
\begin{figure}[htbp]
    \centering
    \resizebox{\columnwidth}{!}{
    \begin{tikzpicture}[scale=0.6]
        % Axes
        \draw[->] (-0.5,0) -- (14,0) node[right] {Time ($T_u$)};
        \draw[->] (0,-0.5) -- (0,2) node[above] {TX Pin Logic};

        % Ticks
        \foreach \x in {1,2,3,4,5,6,7,8,9,10,11,12,13}
            \draw (\x,0.1) -- (\x,-0.1) node[below, scale=0.7] {\x};

        % Logic High/Low labels
        \node[left, scale=0.8] at (0, 1) {HIGH (1)};
        \node[left, scale=0.8] at (0, 0.25) {LOW (0)};

        % Waveform for "A" (.-) and "N" (-.)
        \draw[thick, blue]
            (0,0) -- (0,1) -- (1,1)   % A: dit
            -- (1,0) -- (2,0)         % intra-gap
            -- (2,1) -- (5,1)         % A: dah
            -- (5,0) -- (8,0)         % inter-character gap
            -- (8,1) -- (11,1)        % N: dah
            -- (11,0) -- (12,0)       % intra-gap
            -- (12,1) -- (13,1)       % N: dit
            -- (13,0) -- (14,0);      % return to idle

        % Annotations
        \draw[<->, dashed, red] (0, 1.4) -- (5, 1.4) node[midway, above] {Letter 'A' ( $\cdot$ $\textbf{-}$ )};
        \draw[<->, dashed, red] (8, 1.4) -- (13, 1.4) node[midway, above] {Letter 'N' ( $\textbf{-}$ $\cdot$ )};
        \draw[<->, dashed, green!50!black] (5, 0.5) -- (8, 0.5) node[midway, above, scale=0.8] {3-$T_u$ Gap};

    \end{tikzpicture}
    }
    \caption{Baseband OOK transmission waveform for the word "AN", illustrating the strict 1:3 timing ratio of standard Morse protocol.}
    \label{fig:morse_an}
\end{figure}



\section{Visible Light Baseline}

\subsection{Circuit Implementation}
The baseline system operates entirely in the visible spectrum. 
\begin{itemize}
    \item Transmitter (TX): An Arduino microcontroller drives a standard visible LED. The LED's anode connects to Digital Pin 5, while the cathode is grounded through a $220\Omega$ current-limiting resistor. 
    \item Receiver (RX): The optical sensor is a Light Dependent Resistor (LDR) arranged in a voltage divider. The 5V supply is connected to one leg of the LDR. The secondary leg connects to the Arduino's Analog Pin A0 and a $10\text{k}\Omega$ pull-down resistor connected to ground. 
\end{itemize}

\subsection{Hardware Latency and The "Iceberg Effect"}
When bridging analog optical sensing and digital microcontrollers, signal truncation occurs. As shown in Fig. \ref{fig:iceberg}, an expected "Dah" signal transmitted for 600ms is detected for a fraction of that time. LDRs exhibit notable rise/fall latency, resulting in a trapezoidal curve. To counteract this, our receiver employs widened tolerance windows as listed in Table \ref{tab:morse_timing-modified}.

\begin{figure}[htbp]
    \centering
    \resizebox{\columnwidth}{!}{
    \begin{tikzpicture}[scale=0.8]
        \draw[->] (0,0) -- (8,0) node[right] {Time (ms)};
        \draw[->] (0,0) -- (0,5) node[above] {Analog Value};
        \draw[thick, blue, fill=blue!10] (0,0) -- (1.5,0) -- (2.5,4) -- (5.5,4) -- (7.5,0) -- (8,0);
        \node[blue] at (4, 2) {LDR Response Profile};
        \draw[red, thick, dashed] (0,3.5) -- (8,3.5);
        \node[red, right] at (8,3.5) {High Static Threshold};
        \draw[green!60!black, thick, dashed] (0,1.5) -- (8,1.5);
        \node[green!60!black, right] at (8,1.5) {Ideal Threshold};
        \draw[dotted, red] (2.375, 3.5) -- (2.375, -0.5); 
        \draw[dotted, red] (5.75, 3.5) -- (5.75, -0.5); 
        \draw[dotted, green!60!black] (1.875, 1.5) -- (1.875, -1);   
        \draw[dotted, green!60!black] (6.75, 1.5) -- (6.75, -1);   
        \draw[<->, red, thick] (2.375, -0.5) -- (5.75, -0.5) node[midway, below] {Detected: $\approx$330ms};
        \draw[<->, green!60!black, thick] (1.875, -1) -- (6.75, -1) node[midway, below] {Actual: $\approx$600ms};
    \end{tikzpicture}
    }
    \caption{The "Iceberg Effect" illustrating temporal truncation caused by improper static thresholding of an LDR signal.}
    \label{fig:iceberg}
\end{figure}

\begin{table}[htbp]
\caption{Adaptive Morse Protocol Timings to combat "Iceberg Effect"}
\label{tab:morse_timing-modified}
\centering
\begin{tabular}{|l|c|l|}
\hline
\textbf{Protocol Element} & \textbf{Logic State} & \textbf{Duration} \\ \hline
Dot (Dit)                 & HIGH (ON)            &  $150\text{ ms}$ \\ \hline
Dash (Dah)                & HIGH (ON)            &  $450\text{ ms}$ \\ \hline
Intra-character gap       & LOW (OFF)            &  $350\text{ ms}$ \\ \hline
Inter-character gap       & LOW (OFF)            &  $750\text{ ms}$ \\ \hline
Inter-word gap            & LOW (OFF)            &  $1600\text{ ms}$ \\ \hline
\end{tabular}
\end{table}




\subsection{Visible Light Results}
The baseline system achieved a maximum reliable transmission distance of only 11-12 cm in room light, and 27 cm in total darkness.

\section{Infrared Evolution and SDM}

\subsection{Circuit Implementation}
To resolve the noise and range issues of the baseline VLC system, the physical layer was upgraded to a 940-nm Infrared (IR) ecosystem.
\begin{itemize}
    \item Transmitter (TX): The visible LED is replaced by an IR LED. 
    \item Receiver (RX) Sensor: A high-speed IR photodiode replaces the LDR, operating in reverse-bias mode. The cathode connects to 5V, and the anode connects to Analog Pin A0 and a pull-down load resistor ($R_L$) to ground.
\end{itemize}

\subsection{Hardware Optimization and SDM}
Operating in reverse-bias, the photodiode produces a photocurrent ($I_{pd}$) proportional to the incident photon flux; consequently, the voltage sampled by the ADC is governed by $V_{adc} = I_{pd} \times R_L$. By increasing $R_L$ from $10\text{k}\Omega$ to $1\text{M}\Omega$, sensitivity was maximized, extending the raw range to 1.57 meters indoors.

To isolate transmission pulses from fluctuating indoor IR noise, we developed the Stepwise Derivative Method (SDM). The core challenge is that the "dark" voltage of a photodiode is not zero; it is a floating DC offset that changes as room lights are toggled. To address this, the algorithm first calculates a moving average to capture the slow-moving ambient noise floor. This part of the algorithm functions inherently as a low-pass filter:

\begin{align}
    V_{noise}[n] = \frac{1}{N} \sum_{i=0}^{N-1} V_{adc}[n-i] \label{eq:noise}
\end{align}

The SDM achieves its adaptive high-pass characteristics by subtracting this low-frequency baseline from the raw instantaneous sampled voltage. This derivative-like operation ignores the steady-state DC bias and extracts only the sudden, high-frequency voltage "steps" caused by the transmitter:

\begin{align}
    V_{signal}[n] = V_{adc}[n] - V_{noise}[n] \label{eq:highpass}
\end{align}

The final detection threshold is then defined dynamically above the noise floor:

\begin{align}
    V_{threshold}[n] = V_{noise}[n] + \sigma \label{eq:thresh}
\end{align}
Where $\sigma$ is a static signal margin constant. This ensures the "Detection Zone" floats above the noise floor, maintaining a constant sensitivity regardless of ambient brightness.

\subsection{SDM Initialization and Training Sequence}
Because the rolling buffer requires a stabilization period to accurately map the ambient noise floor, the system mandates an initialization phase prior to data transmission. During this phase, a known training sequence—the phrase "HELLO WORLD"—is transmitted 10 times. 

Based on the timing ratios defined in Table \ref{tab:morse_timing}, a single transmission of "HELLO WORLD" consumes exactly 111 $T_u$ (inclusive of all intra-character, inter-character, and inter-word gaps). Therefore, the full 10-cycle training sequence spans 1110 $T_u$. Given our baseline $T_u = 200\text{ ms}$, this equates to a deliberate 222-second pre-transmission window. This ensures the SDM buffer ($V_{noise}$) fully saturates with representative environmental data, allowing the dynamic noise margin to settle completely before the receiver attempts to decode unknown payload characters.

\section{Frequency Modulation and Software-Defined Filtering}

\subsection{Limitations in Direct Sunlight and Hardware Gain Control}
While the initial IR setup successfully mitigated indoor noise, it typically fails in direct sunlight. The sun acts as a high-intensity IR emitter, causing the photodiode to produce a massive steady-state photocurrent. This threatens to saturate the 10-bit ADC at a constant 1023, effectively "blinding" the receiver and removing the dynamic range necessary for processing.

To prevent "hard" saturation, the receiver's load resistor ($R_L$) is decreased (e.g., from $1\text{M}\Omega$ to $10\text{k}\Omega$) when operating outdoors. While this keeps the DC component within the ADC's measurable range, it significantly reduces the signal-to-noise ratio (SNR), as the intentional AC pulses become minute fluctuations on a massive DC floor. This necessitates a transition to a frequency-domain detection approach to isolate the signal. 

By reducing $R_L$, we trade off raw voltage swing for headroom, relying on the high-gain nature of the Goertzel algorithm to recover the attenuated signal from the resulting narrow dynamic range.

\subsection{2~kHz On-Off Keying (OOK) Modulation}
To resolve solar saturation, we transition to a Software-Defined Optical paradigm. Instead of a static DC pulse, the transmitter modulates the IR LED with a 2~kHz square wave carrier during every Morse "mark" (dit or dah).

This process, acting as a form of subcarrier modulation, involves gating the continuous DC transmission pulse with a high-frequency sequence. By toggling the LED at 2~kHz (a 50\% duty cycle), we effectively perform a frequency-domain shift:
\begin{itemize}
    \item \textbf{DC Offset (Sun):} The solar interference resides at 0Hz.
    \item \textbf{Data Signal:} The intentional Morse pulses are now "nested" inside a 2000Hz carrier.
\end{itemize}
This separation allows the receiver to ignore the saturated 0Hz baseline and focus entirely on the energy oscillating at the target frequency band. This is implemented in software using the \texttt{tone()} function, which leverages the microcontroller's internal timers to generate a precise square wave. \\
Furthermore, operating the LED at a 50\% duty cycle reduces the active transmission power consumption by half compared to the continuous DC pulses observed during the Baseband IR, highly benefiting battery-constrained remote sensor nodes.

\begin{figure}[htbp]
    \centering
    \begin{tikzpicture}[scale=1]
        % Axes
        \draw[->] (0,0) -- (6,0) node[right] {Time};
        \draw[->] (0,-0.5) -- (0,1.5) node[above] {Amplitude};
        
        % Logic Envelope (Dashed)
        \draw[red, thick] (0,0) -- (1,0) -- (1,1) -- (3,1) -- (3,0) -- (4,0) -- (4,1) -- (5,1) -- (5,0);
        \node[red, scale=0.8] at (2,1.2) {Morse 'A' Envelope (.-)};

        % 2~kHz Square Wave Modulated Carrier (Dah)
        \draw[blue] (1,0) -- (1.1,0) -- (1.1,1) -- (1.2,1) -- (1.2,0) -- (1.3,0) -- (1.3,1) -- (1.4,1) -- (1.4,0) -- (1.5,0) -- (1.5,1) -- (1.6,1) -- (1.6,0) -- (1.7,0) -- (1.7,1) -- (1.8,1) -- (1.8,0) -- (1.9,0) -- (1.9,1) -- (2.0,1) -- (2.0,0) -- (2.1,0) -- (2.1,1) -- (2.2,1) -- (2.2,0) -- (2.3,0) -- (2.3,1) -- (2.4,1) -- (2.4,0) -- (2.5,0) -- (2.5,1) -- (2.6,1) -- (2.6,0) -- (2.7,0) -- (2.7,1) -- (2.8,1) -- (2.8,0) -- (2.9,0) -- (2.9,1) -- (3.0,1) -- (3.0,0);
        
        % 2~kHz Square Wave Modulated Carrier (Dit)
        \draw[blue] (4,0) -- (4.1,0) -- (4.1,1) -- (4.2,1) -- (4.2,0) -- (4.3,0) -- (4.3,1) -- (4.4,1) -- (4.4,0) -- (4.5,0) -- (4.5,1) -- (4.6,1) -- (4.6,0) -- (4.7,0) -- (4.7,1) -- (4.8,1) -- (4.8,0) -- (4.9,0) -- (4.9,1) -- (5.0,1) -- (5.0,0);
        
        \node[blue, align=center, scale=0.8] at (4.5,1.5) {2~kHz Square Wave\\Carrier};
        \node[red, align=center, scale=0.8] at (3,-0.3) {0Hz Solar Bias Ignored};
    \end{tikzpicture}
    \caption{Modulation: The logic envelope acts as a gate for the 2~kHz square wave. This subcarrier modulation enables frequency-selective filtering at the receiver.}
    \label{fig:modulation}
\end{figure}

\subsection{Sampling and Goertzel Digital Signal Processing}
To detect the 2~kHz carrier without external hardware filters, the receiver's ADC must satisfy the Nyquist theorem (requiring a minimum sampling rate of 4kHz). We modify the \texttt{ADCSRA} register to change the prescaler from 128 to 16, accelerating the sampling rate from $\approx 10\text{kHz}$ to $\approx 60\text{kHz}$, which provides massive oversampling for high-resolution frequency detection.

We then implement the Goertzel algorithm \cite{smith1997} to act as a digital bandpass filter. Unlike a full Fast Fourier Transform (FFT), the Goertzel algorithm is computationally optimized to calculate the magnitude of a single target frequency ($f_{target} = 2000\text{Hz}$), making it ideal for the limited memory of microcontrollers \cite{mitra2006}. 

During the real-time loop, the system updates a second-order Infinite Impulse Response (IIR) filter state:
\begin{equation}
    s[n] = x[n] + \left(2 \cos\left(\frac{2\pi k}{N}\right) \cdot s[n-1]\right) - s[n-2]
\end{equation}
After $N=80$ samples (which takes approximately $1.33\text{ ms}$ to process at 60~kHz), the squared magnitude $M^2$ is evaluated. Because DC sunlight (0Hz) and 120Hz indoor flicker do not resonate with the pre-computed coefficient, their resultant magnitude approaches zero. This effectively neutralizes the moving noise floor that required adaptive SDM in IR, allowing Frequency-Modulated IR to utilize a highly reliable static magnitude threshold to "see" the 2~kHz flashes even through a saturated DC bias.

The efficacy of this software-defined approach is evidenced by the raw sampled data. In direct sunlight, the ambient DC floor may occupy $\approx 90\%$ of the ADC range (e.g., a baseline of $\approx 950$), while the 2~kHz carrier oscillates with a peak-to-peak amplitude of only $\pm 15$ units. Traditional thresholding (SDM) fails in this context because environmental noise fluctuations often exceed the signal amplitude. However, because the Goertzel algorithm concentrates the signal energy into a single frequency bin, it produces a high-magnitude resonance that is easily distinguishable from the DC solar offset and 120Hz artificial lighting flicker, allowing for a reliable static magnitude threshold.

\section{Experimental Results}

\subsection{Throughput and Bitrate Analysis}
We evaluated the efficiency of the three methods by transmitting a 44-byte (352-bit) test sentence. While the VLC setup required slower custom timings (101.8s) to account for photo-conductive latency, the transition to standard Morse timing in IR and Frequency-Modulated IR communication significantly improved throughput. Because Morse code utilizes variable-length OOK encoding, instantaneous throughput fluctuates based on character transmission length. A 44-byte pangram was explicitly chosen as the test payload to force the transmission of all alphabet symbol variants, thereby providing a normalized, worst-case representative average for system bitrate.

\begin{table}[htbp]
\caption{System Performance Comparison (44-Byte Sentence)}
\label{tab:bps}
\centering
\begin{tabular}{|l|c|c|l|}
\hline
\textbf{Receiver Mode of } & \textbf{Time (s)} & \textbf{Bitrate (bps)} & \textbf{Environment} \\ 
\textbf{Communication} & & \\ \hline
Visible Light & 101.8 & 3.46 & Dark/Indoor only \\ \hline
Baseband IR  & 81.4 & 4.32 & Indoor (Non-Solar) \\ \hline
Frequency-Modulated IR & 81.4 & 4.32 & \textbf{Direct Sunlight} \\ \hline
\end{tabular}
\end{table}

As shown in Table \ref{tab:bps}, Frequency-Modulated IR maintains the peak bitrate of 4.32 bps even in high-solar environments where VLC and Baseband IR systems experience complete signal saturation and $<10$\% decoding accuracy.

\subsection{Range Analysis}
The integration of impedance-tuned hardware with software-defined filtering yielded profound improvements. As illustrated in Fig. \ref{fig:graph}, the visible light baseline failed abruptly at 27 cm. The unoptimized IR setup ($10\text{k}\Omega$) reached 47 cm. The fully optimized IR-SDM setup ($1\text{M}\Omega$) successfully maintained 100\% decoding accuracy up to 157 cm indoors. Decoding accuracy was empirically measured as the percentage of correctly reconstructed characters from the 44-byte test payload at varying distances.

\begin{figure}[htbp]
    \centering
    \resizebox{0.9\columnwidth}{!}{
    \begin{tikzpicture}
        \begin{axis}[
            xlabel={Transmission Distance (cm)},
            ylabel={Decoding Accuracy (\%)},
            xmin=0, xmax=180,
            ymin=0, ymax=110,
            legend pos=south west,
            grid=both,
            mark size=2.5pt
        ]
        \addplot[color=red, mark=square*] coordinates {(0,100) (15,100) (27,100) (30,0)}; 
        \addplot[color=blue, mark=triangle*] coordinates {(0,100) (30,100) (47,100) (52,0)}; 
        \addplot[color=green!60!black, mark=*] coordinates {(0,100) (100,100) (157,100) (165,0)}; 
        \legend{Visible (LDR), IR ($10\text{k}\Omega$), IR ($1\text{M}\Omega$)}
        \end{axis}
    \end{tikzpicture}
    }
    \caption{Comparative analysis of decoding accuracy over distance across the configurations.}
    \label{fig:graph}
\end{figure}

% \section{Conclusion and Future Work}
% This paper presented an evolutionary approach to overcoming the physical layer limitations of low-cost optical wireless communication. By systematically advancing from a basic visible light LDR circuit to an Infrared Stepwise Derivative Method (SDM), we mitigated indoor thermal noise and extended the range up to 2m. The implementation of a software-defined Goertzel filter on the 8-bit ATmega328P architecture proved to be the most robust solution. By shifting from hardware-reliant IR sensing to digital signal processing, the system successfully bypassed the 0Hz DC saturation caused by direct sunlight. This "Virtual Filter" approach allows low-cost microcontrollers to achieve high-selectivity communication that traditionally requires expensive specialized hardware. Future work will focus on optimizing the Goertzel block size ($N$) to further increase the bitrate while maintaining the noise-rejection benefits of the software-defined optical link.

\section{Conclusion and Future Work}
This paper demonstrated a systematic progression in overcoming the physical layer limitations of low-cost optical wireless communication. By evolving from basic visible light sensing to an adaptive Infrared system, we mitigated photo-conductive latency and indoor thermal noise. By shifting the complexity from hardware to a software-defined paradigm—utilizing 2~kHz OOK modulation and an optimized Goertzel digital filter—the system successfully maintained 100\% decoding accuracy at ranges up to 2m, even under direct solar interference. This "virtual filter" approach highlights the potential for low-power microcontrollers to perform sophisticated signal recovery in high-noise environments without requiring specialized commercial hardware. Future work will explore increasing the baud rate by optimizing the Goertzel block size $N$ for higher-speed telemetry.

\bibliographystyle{IEEEtran}
\bibliography{references}

\appendices
\section{Visible Light Source Code}

\section{Infrared SDM Implementation}

\section{Software-Defined High-Solar Ecosystem}

\end{document}